`r opts_chunk$set(cache=TRUE, tidy=FALSE)`
# Example 2 from Rossel's Paper on lavaan
```{r setup, message=FALSE}
library(lavaan)
Data <- PoliticalDemocracy
```

This example is an elaboration on Example 2 from Yves Rossel's Journal of Statistical Software Article (see [here](http://www.jstatsoft.org/v48/i02/paper)).

## M0: Basic Measurement model
```{r basic_measurement_model}
m0_model <- '
# measurement model
ind60 =~ x1 + x2 + x3
dem60 =~ y1 + y2 + y3 + y4
dem65 =~ y5 + y6 + y7 + y8
'

m0_fit <- cfa(m0_model, data=Data)
```

* `m0` defines a basic measurement model that permits correlated factors.  Note that it does not have correlations between corresponding democracy indicator measures over time.

**Questions:**

* Is it a good model?

```{r m0_fit_measures}
fitmeasures(m0_fit)
```

* cfi suggests a reasonable model, but RMSEA is quite large.

```{r m0_standardised_parameters}
inspect(m0_fit, 'standardized')
```

* The table of standardised loadings show all factor loadings to be large.

```{r m0_mod_indices}
m0_mod <- modificationindices(m0_fit)
head(m0_mod[order(m0_mod$mi, decreasing=TRUE), ], 12)
```

* The table of largest modification indices suggest a range of ways that the model could be improved. Because the sample size is small, particular caution needs to be taken with these.
* Several of these modifications concern the expected requirement to permit indicator variables at different time points to correlate (e.g., `y2` with `y6`, `y3` with `y7`).
* It may also be that some pairs of items are correlated more than others. For example, the following correlation matrix shows how `y6` and `y8` have a particularly large correlation.

```{r}
round(cor(Data[,c('y5', 'y6', 'y7', 'y8')]), 2)
```


* What are the correlations between the factors?

```{r}
cov2cor(inspect(m0_fit, "coefficients")$psi)
```

This certainly suggests that factors are strongly related, especially the two demographics measures.


## M1: Correlated item measurement model
This next model permits corresponding democracy measures from the two points to be correlated.

```{r correlated_measurement_model}
m1_model <- '
    # measurement model
    ind60 =~ x1 + x2 + x3
    dem60 =~ y1 + y2 + y3 + y4
    dem65 =~ y5 + y6 + y7 + y8
    
    # correlated residuals
    y1 ~~ y5
    y2 ~~ y6
    y3 ~~ y7
    y4 ~~ y8
'

m1_fit <- cfa(m1_model, data=Data)
```

* Is this an improvement over `m0` with uncorrelated indicators?
* Does `m1` have good fit in and of itself?

```{r}
anova(m0_fit, m1_fit)
round(cbind(m0=inspect(m0_fit, 'fit.measures'), 
            m1=inspect(m1_fit, 'fit.measures')), 3)
```

* It is a significant improvement. 
* RMSEA and other fit measurs are substantially improved.
* The relatively small sample size makes it somewhat difficult to see how much further improvements should continue. In general, the RMSEA suggests that further improvements are possible but it may be less clear on how to proceed in a principled way.




# M2: Basic SEM
```{r m2_model}
m2_model <- '
    # measurement model
    ind60 =~ x1 + x2 + x3
    dem60 =~ y1 + y2 + y3 + y4
    dem65 =~ y5 + y6 + y7 + y8
    
    # correlated residuals
    y1 ~~ y5
    y2 ~~ y6
    y3 ~~ y7
    y4 ~~ y8

    # regressions
    dem60 ~ ind60
    dem65 ~ ind60 + dem60
'

m2_fit <- sem(m2_model, data=Data)
```

* Is fit the same as model 1 as I would expect?

```{r m2_chi_square_check}
rbind(m1 = fitMeasures(m1_fit)[c('chisq', 'rmsea')], 
    m2 = fitMeasures(m2_fit)[c('chisq', 'rmsea')])
```
Yes, it is.

* Assuming democracy 1965 is the depenent variable, how can we get the information typically available in multiple regression output?
    * R-squared?
    * Unstandardised regression coefficients?
    * Standardised regression coefficients?
    * Standard errors, p-values,  and confidence intervals on unstandardised coefficients?

```{r}
# m2_fit <- sem(m2_model, data=Data)

# r-square for dem-65
inspect(m2_fit, 'r2')['dem65']

# Unstandardised regression coefficients
inspect(m2_fit, 'coef')$beta['dem65', ]

# Standardised regression coefficients
subset(inspect(m2_fit, 'standardized'), lhs == 'dem65' & op == '~')

# Just a guess, may not be correct:
# coefs <- data.frame(coef=inspect(m2_fit, 'coef')$beta['dem65', ],
#       se=inspect(m2_fit, 'se')$beta['dem65', ])
# coefs$low95ci <- coefs$coef - coefs$se * 1.96
# coefs$high95ci <- coefs$coef + coefs$se * 1.96
```



