# Example 2 from the Rossel's Paper on lavaan
```{r}
library(lavaan)
Data <- PoliticalDemocracy
```

This example is an elaboration on Example 2 from Yves Rossel's Journal of Statistical Software Article (see [here](http://www.jstatsoft.org/v48/i02/paper)).

# Basic SEM
```{r}
m0_model <- '
# measurement model
ind60 =~ x1 + x2 + x3
dem60 =~ y1 + y2 + y3 + y4
dem65 =~ y5 + y6 + y7 + y8
'

m0_fit <- cfa(m1_model, data=Data)
```

* `m0` defines a basic measurement model that permits correlated factors. Note that it does not have correlations between corresponding democracy indicator measures over time.

Questions:

* Is it a good model?

```{r}
fitmeasures(m0_fit)
```

* cfi suggests a reasonable model, but RMSEA is quite large.

```{r}
inspect(m0_fit, 'standardized')
```

* The table of standardised loadings show all factor loadings to be large.

```{r}
m0_mod <- modificationindices(m0_fit)
head(m0_mod[order(m0_mod$mi, decreasing=TRUE), ], 12)
```

* The table of largest modification indices suggest a range of ways that the model could be improved. Because the sample size is small, particular caution needs to be taken with these.
* Several of these modifications concern the expected requirement to permit indicator variables at different time points to correlate (e.g., `y2` with `y6`, `y3` with `y7`).
* It may also be that 



* What are the correlations between the factors?



```{r}
m1_model <- '
# measurement model
ind60 =~ x1 + x2 + x3
dem60 =~ y1 + y2 + y3 + y4
dem65 =~ y5 + y6 + y7 + y8

# regressions
dem60 ~ ind60
dem65 ~ ind60 + dem60
'

m1_fit <- sem(m1_model, data=Data)
```


```{r}
summary(m1_fit)
inspect(m1_fit, 'standardized')
inspect(m1_fit, 'r2')[c('dem60', 'dem65')]

```

