`r opts_chunk$set(cache=TRUE, tidy=FALSE)`
# Example 2 from Rossel's Paper on lavaan
```{r setup, message=FALSE}
library(lavaan)
Data <- PoliticalDemocracy
```

This example is an elaboration on Example 2 from Yves Rossel's Journal of Statistical Software Article (see [here](http://www.jstatsoft.org/v48/i02/paper)).

# M0: Basic Measurement model
```{r basic_measurement_model}
m0_model <- '
# measurement model
ind60 =~ x1 + x2 + x3
dem60 =~ y1 + y2 + y3 + y4
dem65 =~ y5 + y6 + y7 + y8
'

m0_fit <- cfa(m0_model, data=Data)
```

* `m0` defines a basic measurement model that permits correlated factors.  Note that it does not have correlations between corresponding democracy indicator measures over time.

**Questions:**

* Is it a good model?

```{r m0_fit_measures}
fitmeasures(m0_fit)
```

* cfi suggests a reasonable model, but RMSEA is quite large.

```{r m0_standardised_parameters}
inspect(m0_fit, 'standardized')
```

* The table of standardised loadings show all factor loadings to be large.

```{r m0_mod_indices}
m0_mod <- modificationindices(m0_fit)
head(m0_mod[order(m0_mod$mi, decreasing=TRUE), ], 12)
```

* The table of largest modification indices suggest a range of ways that the model could be improved. Because the sample size is small, particular caution needs to be taken with these.
* Several of these modifications concern the expected requirement to permit indicator variables at different time points to correlate (e.g., `y2` with `y6`, `y3` with `y7`).
* It may also be that some pairs of items are correlated more than others. For example, the following correlation matrix shows how `y6` and `y8` have a particularly large correlation.

```{r}
round(cor(Data[,c('y5', 'y6', 'y7', 'y8')]), 2)
```


* What are the correlations between the factors?

```{r}
cov2cor(inspect(m0_fit, "coefficients")$psi)
```

# M1: Correlated item measurement model
```{r correlated_measurement_model}
m1_model <- '
    # measurement model
    ind60 =~ x1 + x2 + x3
    dem60 =~ y1 + y2 + y3 + y4
    dem65 =~ y5 + y6 + y7 + y8
    
    # correlated residuals
    y1 ~~ y5
    y2 ~~ y6
    y3 ~~ y7
    y4 ~~ y8
'

m1_fit <- cfa(m1_model, data=Data)
```













# M2: ...
```{r m1_basic_sem}
m2_model <- '
# measurement model
ind60 =~ x1 + x2 + x3
dem60 =~ y1 + y2 + y3 + y4
dem65 =~ y5 + y6 + y7 + y8

# regressions
dem60 ~ ind60
dem65 ~ ind60 + dem60
'

m2_fit <- sem(m2_model, data=Data)
```


```{r m1_summary}
#summary(m1_fit)
#inspect(m1_fit, 'standardized')
#inspect(m1_fit, 'r2')[c('dem60', 'dem65')]
```


