`r opts_chunk$set(tidy=FALSE)`
This exercise examines the first example shown in 
<http://www.jstatsoft.org/v48/i02/paper>.
It's a three-factor confirmatory factor analysis example with three items per factor.
All three latent factors are permitted to correlate.

* `x1` to `x3` load on a `visual` factor
* `x4` to `x6` load on a `textual` factor
* `x7` to `x9` load on a `speed` factor

```{r setup, message=FALSE}
library('lavaan')
library('Hmisc')
cases <- HolzingerSwineford1939
```

# Quick examination of data

```{r }
str(cases)
Hmisc::describe(cases)
```

The data set include `r nrow(cases)` observations. It includes a few demographic variables (e.g., sex, age in years and months, school, and grade). It includes nine variables  that are the observed test scores used in the subsequent CFA.

# Fit CFA
```{r fit_m1}
m1_model <- ' visual  =~ x1 + x2 + x3
              textual =~ x4 + x5 + x6
              speed   =~ x7 + x8 + x9
'

m1_fit <- cfa(m1_model, data=cases)

```

* model syntax is specified as a character variable
* `cfa` is one model fitting function in `lavaan`. The command includes many options. Data can be specified as a data frame, as it is here using the `data` argument. Alternatively covariance matrix, vector of means, and sample size can be specified.
* From what I can tell, `lavaan` is the parent model fitting function that can take a `model.type` argument of `'cfa'`, `'sem'`, or `'growth'`. Thus, the arguments to `model.type` are functions themselves which just call `lavaan` with particular argument values.

## Show parameter table
```{r}
parTable(m1_fit)
```

* What do the columns mean?
    * `id`: numeric identifier for the parameter
    * `lhs`: left hand side variable name
    * `op`: operator (see page 7 of http://www.jstatsoft.org/v48/i02/paper); `=~` 
    is manifested by; `~~` is correlated with.
    * `rhs`: right hand side variable name
    * `user`: 1 if parameter was specified by the user, 0 otherwise
    * `group`: presumably used in multiple group analysis
    * `free`: Nonzero elements are free parameters in the model
    * `ustart`: The value specified for fixed parameters
    * `exo`: ???
    * `label`: Probably just an optional label???
    * `eq.id`: ??? 
    * `unco`: ???


* The model syntax used in `lavaan` incorporates a lot of parameters by default to permit a tidy model syntax. The exact nature of these parameters is also determined by options in the `cfa`, `sem` and other model fitting fucntions.
* `parTable` is a method

It shows that the latent factors are allowed to intercorrelate. The `cfa` function has an an argument `orthogonal`. It defaults to FALSE which permits correlated factors.

```{r}
parTable(cfa(m1_model, data=cases, orthogonal=TRUE))[22:24, ]
```
When `orthogonal=TRUE` is specified, the covariance of latent factors is constrained to zero. This is reflected in `free=0` (i.e., it's not free to vary) and `ustart=0` (the constrained value is zero) in the parameter table.

Returning to the original parameter table:

* Variance parameters (`op=~~` where `lhs` is the same as `rhs`) are included for all observed and latent variables.

## Summarise fit
```{r}
summary(m1_fit)
```

The default `summary` method shows $\chi^2$, $df$, p-value for the overall model, unstandardised parameter estimates, in some cases with significance tests.


## Getting fit statistics
There are multiple ways of getting fit statistics

```{r}
fitMeasures(m1_fit)
# equivalent to:
# inspect(m1_fit, 'fit.measures')

fitMeasures(m1_fit)['rmsea']
fitMeasures(m1_fit, c('rmsea', 'rmsea.ci.lower', 'rmsea.ci.upper'))


summary(m1_fit, fit.measures=TRUE)
```


* I assume that lavaan uses S4 classes which makes extracting elements a little different to S3 classes.
* The above code shows how to extract fit measures.
* While it is not clear hear, it appears that `rmsea.ci.lower` and `rmsea.ci.upper` refer to 90% lower and upper confidence intervals.
* Adding `fit.measures=TRUE` provides a way of displaying 


## Modification indices
```{r}
m1_mod <- modificationIndices(m1_fit)
head(m1_mod[order(m1_mod$mi, decreasing=TRUE), ], 10)
```

* The `modificationIndices` function returns modification indices and expected parameter changes (EPCs). 
* The second line above sorts the rows of the modification indices table in decreasing order and shows those parameters with the 10 largest values.

```{r}
m2_model <- ' visual  =~ x1 + x2 + x3 + x9
              textual =~ x4 + x5 + x6
              speed   =~ x7 + x8 + x9
'

m2_fit <- cfa(m2_model, data=cases)
anova(m1_fit, m2_fit)
```

* Note that more than one empty line at the end of the model definition seems to cause an error.
* TODO: Work out why the change in $\chi^2$ 
`r anova(m1_fit, m2_fit)[2, 'Chisq diff']`
is different to the value of the modification index
`r m1_mod[m1_mod$lhs == 'visual' &  m1_mod$rhs == 'x9', 'mi']`.

## Standardised parameters
```{r}
summary(m1_fit)
standardizedSolution(m1_fit)
```



