`r opts_chunk$set(cache=TRUE, tidy=FALSE)`

# Path Analysis Example
```{r, message=FALSE}
library(psych)
library(lavaan)

```


## Simulate data
Let's simulate some data:

* three orthogonal predictor variables
* one mediator variable
* one dependent variable

```{r}
set.seed(1234)
N <- 1000
iv1 <- rnorm(N, 0, 1)
iv2 <- rnorm(N, 0, 1)
iv3 <- rnorm(N, 0, 1)
mv <- rnorm(N, .2 * iv1 + -.2 * iv2 + .3 * iv3, 1)
dv <- rnorm(N, .8 * mv, 1)
data_1 <- data.frame(iv1, iv2, iv3, mv, dv)
```

## Traditional examination of dataset
* Is a regression consistent with the model?

```{r}
summary(lm(mv ~ iv1 + iv2 + iv3, data_1))
```

This is broadly similar to the equation predicting `mv`.

```{r}
summary(lm(dv ~ iv1 + iv2 + iv3 + mv, data_1))
```
Given that the simulation is based on complete mediation, the true regression coefficients for the ivs are zero. The results of the multiple regression predicting the `dv` from the `iv`s and `mv` is consistent with this.

What are the basic descriptive statistics and intercorrelations?

```{r}
psych::describe(data_1)
pairs.panels(data_1, pch='.')
```


## M1 Fit Path Analysis model

```{r}
m1_model <- '
dv ~ mv
mv ~ iv1 + iv2 + iv3
'

m1_fit <- sem(m1_model, data=data_1)
```

Are the regression coefficients the same?

```{r}
parameterestimates(m1_fit)
```

All the coefficients are in the ball park of what is expected.

Does the model provide a good fit?
```{r}
fitmeasures(m1_fit)
```

* The fitted model should provide a good fit because the fitted model is identical to the model used to simulate the data.
* In this case, the p-value and the fit measures are consistent with the data being generated from the model specified.


## Calculate and test indirect effects
```{r}
m2_model <- '
    dv ~ b1*mv
    mv ~ a1*iv1 + a2*iv2 + a3*iv3
    
    # indirect effects
    iv1_mv := a1*b1
    iv2_mv := a2*b1
    iv3_mv := a3*b1
'

m2_fit <- sem(m2_model, data=data_1)
```

* Note that I needed to name effects before I could define the indirect effect as the product of two effects using `:=` notation.


```{r}
parameterestimates(m2_fit, standardize=TRUE)

```

The above output provide a significance test, and confidence intervals for the indirect effects, and includes standardised effects.



